Adversarial deep neural networks remove nonlinear batch effects from gene-expression data

### 2019-07-16 thesis.tex
Revised abstract

\title{Adversarial Deep Neural Networks Effectively Remove Nonlinear Batch Effects from Gene-Expression Data}
\author{Jonathan Bryan Dayton}
\author[1]{Jonathan B. Dayton}
\author[1]{Stephen R. Piccolo}
\affil[1]{Department of Biology, Brigham Young University, Provo, UT 84602 USA}

@@

\begin{abstract}
	Gene-expression profiling enables researchers to quantify transcription levels in cells, thus providing insight into functional mechanisms of diseases and other biological processes.
	However, because of the high dimensionality of these data and the sensitivity of measuring equipment, expression data often contains unwanted confounding effects that can skew analysis.
	For example, collecting data in multiple runs causes nontrivial differences in the data (known as batch effects),
	known covariates that are not of interest to the study may have strong effects,
	and there may be large systemic effects when integrating multiple expression datasets.
	Additionally, many of these confounding effects represent higher-order interactions that may not be removable using existing techniques that identify linear patterns.
	We created Confounded to remove these effects from expression data.
	Confounded is an adversarial variational autoencoder that removes confounding effects while minimizing the amount of change to the input data.
	We tested the model on artificially constructed data and commonly used gene expression datasets and compared against other common batch adjustment algorithms.
	We also applied the model to remove cancer-type-specific signal from a pan-cancer expression dataset.
	Our software is publicly available at \url{https://github.com/jdayton3/Confounded}.
\end{abstract}

>>>>

\title{Adversarial deep neural networks remove nonlinear batch effects from gene-expression data}
\author{Jonathan Bryan Dayton}
\author[1]{Jonathan B. Dayton}
\author[1]{Stephen R. Piccolo}
\affil[1]{Department of Biology, Brigham Young University, Provo, UT 84602 USA}
%[@Please list my contact information. I don't remember how to do this in LaTeX. But please make sure it looks similar to our previous paper.]

@@

\begin{abstract}
	Gene-expression profiling enables researchers to quantify transcription levels in cells, thus providing insights into functional mechanisms of diseases and other biological processes.
	However, because of the high dimensionality of these data and the sensitivity of measuring equipment, expression data often contains unwanted confounding effects that can skew analyses.
	For example, collecting data in multiple runs can cause nontrivial differences in the data known as batch effects.
	Covariates that are not of interest to the study may have strong effects on gene expression,
	and there may be systemic platform effects when integrating multiple datasets.
	These confounding effects may be driven by higher-order interactions that are not removable using existing techniques that identify linear patterns.
	We created *Confounded* to remove these effects from expression data.
	[@Please modify the following sentence to better reflect the dual-objective aspect and the fact that it uses neural networks.@]
	Confounded is an adversarial variational autoencoder that removes confounding effects while minimizing the amount of change to the input data.
	We tested the model on artificially constructed data and three gene-expression datasets and compared against other batch-adjustment algorithms.
	Our software is available at \url{https://github.com/jdayton3/Confounded}.
\end{abstract}
